---
title: "tast4_rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load libraries

```{r load_libraries}
library(here)
library(sf)
library(tidyverse)
library(janitor)
library(countrycode)
library(tmap)
library(tmaptools)
```

## read in data

Located within the data folder:

-   HDI data from:<http://hdr.undp.org/en/indicators/68606#>

-   shapefile from:<https://hub.arcgis.com/datasets/2b93b06dc0dc4e809d3c8db5cb96ba69_0/explore?location=-2.688200%2C0.000000%2C1.41>

```{r}
HDI <- read_csv(here("data",
                        "Gender Inequality Index.csv"),
                   locale = locale(encoding = "latin1"),
                   na="..")

countries <- st_read(here("data",
                          "World_Countries_(Generalized)",
                          "World_Countries__Generalized_.shp"))
```

## data wrangling

select columns we want and change country name to country code, using the "countrycode" package

we could also just join it without this on the country name column

```{r}
HDIcol <- HDI %>% 
  clean_names() %>% 
  select(country,x2010,x2019) %>% 
  mutate(difference=x2019-x2010) %>% 
  slice(1:189,) %>% 
  mutate(iso_code=countrycode(country,origin='country.name',destination='iso2c'))

```

## rename x2019
```{r}
HDIcol <- HDIcol %>% 
  dplyr::rename("2019_gender_inquality_index"=x2019)
```


## join the data

join the csv to world shape file


```{r}
## join the data
HDI_country <- countries %>% 
  clean_names() %>% 
  left_join(.,
            HDIcol,
            # small idlands have been considered in the country value not their own value 
            by=c("aff_iso"="iso_code")) 
```

## map making

```{r}
HDI_country2 <- HDI_country %>% 
  mutate(summary=case_when(difference>0 ~ "improved",
                           difference=0 ~ "the same",
                           TRUE ~ "worse"))
```
## inspect the map with qtm

```{r}
tmap_mode="plot"
qtm(HDI_country,fill="difference")
```
## mapping 

plan to make four quadrant maps

map1 for 2019_gender_inequality_index

map2 for 2010_gender_inequality_index

map3 for difference between 2019 and 2010 gender inequality index

map4 for legend


```{r}
breaks = c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)
diff_breaks = c(-0.4, -0.3, -0.2, -0.1, 0.0, 0.1)

HDI_country <- HDI_country %>% 
  st_transform(.,crs = "+proj=moll")

# plot each map
tm1 <- tm_shape(HDI_country) + 
  tm_polygons("2019_gender_inequality_index", 
              breaks=breaks,
              palette="PuBu",
              border.alpha = 0.5)+
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(a)", position=c(0,0.85), size=1.5)

tm2 <- tm_shape(HDI_country) + 
  tm_polygons("x2010",
              breaks=breaks, 
              palette="PuBu",
              border.alpha = 0.5) + 
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(b)", position=c(0,0.85), size=1.5)

tm3 <- tm_shape(HDI_country) + 
  tm_polygons("difference",
              breaks=diff_breaks, 
              palette="-Blues",
              midpoint=NA,
              border.alpha = 0.5) + 
  tm_legend(show=FALSE)+
  tm_layout(frame=FALSE)+
  tm_credits("(c)", position=c(0,0.85), size=1.5)

legend <- tm_shape(HDI_country) +
    tm_polygons("2019_gender_inequality_index",
                breaks=breaks,
                palette="-PuBu") +
    tm_legend(show=TRUE)+
    tm_layout(legend.only = TRUE, legend.position=c(0.3,0.25),asp=0.1)+
    tm_shape(HDI_country) +
    tm_polygons("difference",
                palette="-Blues",
                midpoint=NA) +
    # tm_scale_bar(position=c(0.2,0.04), text.size=0.6)+
    tm_compass(north=0, position=c(0.6,0.6))+
    tm_layout(legend.only = TRUE, legend.position=c(0.1,0.1),asp=0.1)+
    tm_credits("Mapped data:\nUN Gender Inequality Index\nWorld outline:\nArcGIS Hub",
               position=c(0.3,0.2),just = "left")
  
t=tmap_arrange(tm1, tm2, tm3, legend, ncol=2)

t
```

